---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Set working directory to a local Github location for RepData_PeerAssessment1.
```{r workingdir, echo=TRUE}
setwd("C:/Users/NNarasim/Git/RepData_PeerAssessment1")
library(data.table)
```

## Loading and preprocessing the data
This reads the contents of the CSV file and removes all NA records.
```{r loaddata, echo=TRUE}
unzip ("activity.zip")
activity <- read.csv("activity.csv")
activity <- activity[complete.cases(activity),]
```



## What is mean total number of steps taken per day?
Aggregate the steps for each day to display in a histogram.
```{r hist, echo=TRUE}
activity_steps <- aggregate(steps ~ date, activity, sum)
hist(activity_steps$steps, xlab = "Total steps per day", ylab = "Number of occurences",
     main = "Total number of steps per day")
act_mean <- mean(activity_steps$steps)
act_mean
act_med <- median(activity_steps$steps)
act_med
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
